---
title: "Simulation_assignment1"
author: "Paulina von Stackelberg"
date: "17 9 2020"
output: html_document
---

In the following chunks, you can find the code for the MCMC simulation for assignment 1.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r load packages, clean workspace, set seed}

rm(list=ls())
set.seed(12)

library(tidyverse)
library(ggplot2)
library(DT)

```


```{r sample the values}
# sample 100 samples from standard normal

samples <- matrix(0, nrow = 5000, 100)
for (i in 1:100) {
  samples[,i]<-rnorm(5000, mean = 0, sd = 1)} # sample 10000 values, 100 times
```




```{r compute the bias, SE, lower and upper bound of 95% CI}

# compute mean

n <- 5000
standard_error <- function(x) 1 / sqrt(length(x)) # Create own function

means <- as.matrix(apply(samples, 2, mean, na.rm=TRUE)) # get the means
stand_error <- as.matrix(apply(samples, 2, FUN = standard_error)) # the population SD is 1 as we are sampling from the standard normal
conf_int_low <- means - 1.96 * (1/sqrt(n)) # calculate the CI
conf_int_high <- means + 1.96 * (1/sqrt(n)) # calculate the CI
bias_calc <- means

# bind together

comb_vals_df <- as.data.frame(cbind(means, stand_error, bias_calc, conf_int_low, conf_int_high))


# change the column names

names(comb_vals_df) <- c("means", "stand_error", "bias_calc", "conf_int_low", "conf_int_high")



# check with conf intervals include 0

comb_vals_df <- comb_vals_df %>%
  as.data.frame() %>%
  mutate(Covered = conf_int_low < 0 & 0 < conf_int_high)


colMeans(comb_vals_df)

```


Create a plot that demonstrates the following:

“A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)

```{r plotting}

limits <- aes(ymax = comb_vals_df$conf_int_high, ymin = comb_vals_df$conf_int_low)

ggplot(comb_vals_df, aes(y=means, x=1:100, colour = Covered)) + 
  geom_hline(aes(yintercept = 0), color = "grey", size = 1) + 
  geom_pointrange(limits) + 
  ylab("Means and 95% CIs")+
  xlab("Simulations")


```



Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.

```{r}
extracted_noncover <- comb_vals_df %>% filter(Covered == FALSE)

# make a nice table using DT

datatable(extracted_noncover)
```



