---
title: "Markup Languages and Reproducible Programming in Statistics"
subtitle: Week 1 Exercise
author:
    - "[Ihnwhi Heo](https://ihnwhiheo.github.io/)"
    - Department of Methodology and Statistics, Utrecht University
date: 'Last modified: `r Sys.setlocale("LC_TIME", "English"); format(Sys.time(), "%d
  %B %Y")`'
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

```{r}
# Load the required packages
library(ggplot2)
library(extrafont)
library(tidyverse)
```

## 1. Sample 100 samples from a standard normal distribution.

```{r}
# Set a seed value for reproducibility
set.seed(322)

# Sample 100 samples from a standard normal distribution
samples <- rnorm(100)
print(samples)
```

## 2. For each of these samples, calculate the following statistics for the mean:

### (1) absolute bias

```{r}
# Calculate the absolute bias
ab.samples <- abs(mean(samples) - 0)
print(ab.samples)
```

### (2) standard error

```{r}
# Make a function that returns a standard error
se.ihnwhi <- function(x) {
  x <- na.omit(x)
  n.x <- length(x)
  sd.x <- sd(x)
  se.x <- sd.x/sqrt(n.x)
  return(se.x)
}

se.samples <- se.ihnwhi(samples)
print(se.samples)
```

### (3) lower bound of the 95% confidence interval

```{r}
lb.samples <- mean(samples) - qnorm(.975)*se.samples
print(lb.samples)
```

### (4) upper bound of the 95% confidence interval

```{r}
ub.samples <- mean(samples) + qnorm(.975)*se.samples
print(ub.samples)
```

## 3. Create a plot that demonstrates the following:

> “A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934).

```{r}
# Set a seed value for replicability
set.seed(322)

# Make a function that creates a confidence interval
ci.ihnwhi <- function(x){
  x <- na.omit(x)
  mean.x <- mean(x)
  n.x <- length(x)
  sd.x <- sd(x)
  se.x <- sd.x/sqrt(n.x)  
  lb.x <- mean.x - qnorm(.975)*se.x
  ub.x <- mean.x + qnorm(.975)*se.x
  info <- c(lb.x, mean.x, ub.x)
  return(info)
}

# Replicate drawing 100 random values 100 times
sim.ci <- replicate(100, rnorm(100), simplify = FALSE)

# Create an empty matrix to store the values
matrix.ci <- matrix(NA, nrow=100, ncol=3)

# Create 100 confidence intervals for replicated datasets
for (i in 1:100) {
  matrix.ci[i,] <- ci.ihnwhi(sim.ci[[i]])
}

# Designate column names for the confidence interval matrix
colnames(matrix.ci) <- c("lower", "mean", "upper")

# Treat matrix.ci as a data frame
matrix.ci <- data.frame(matrix.ci)

ggplot(matrix.ci, aes(x = 1:100, y = min(matrix.ci):max(matrix.ci))) + 
  theme_bw() +
  geom_hline(aes(yintercept=0), color = "blue", size = 1.5) +
  geom_pointrange(aes(ymax = upper, y = mean, ymin = lower), shape=20) +
  xlab("Repliaction 100 times") +
  ylab("95% Confidence Intervals") +
  ggtitle("Simulation Study for 95% Confidence Interval") +
  theme(text=element_text(family="Candara"),
        plot.title = element_text(size = 15, color="black", hjust=0.5, face="bold"))
```

## 4. Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.

```{r}
# Filter rows that does not contain the population values
matrix.ci.condition <- filter(matrix.ci, lower > 0 | upper < 0)

# Print the table that meets the condition
print(matrix.ci.condition)
```
