---
title: "Assignment2_Zhenwei"
author: "Zhenwei Yang"
date: "9/30/2020"
output: html_document
  
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 20px;
  color: DarkGreen;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 16px;
}
h3 { /* Header 3 */
  font-size: 14px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>

---

# 1 Aim

This document demonstrates a simulation study that focuses on survival analysis. The goal is to find how treatment usage would affect the survival probability at the middle of the follow-up. 

# 2 Set up

## 2.1 Notation

This study would simulate 100 participants for 50 time intervals and then replicate for 1000 times. The longitudinanl setting is illustrated as below: 

Let's assume that one predictor that can explain the survival time, denoted as P and treatment usage, deonted as T. Note that T is a time-varying variable which means the status will be changeable at different time points while P is time-invariant. At each time point, the occurance of death as well as treatment follows a logistic distribuiton: 

$$\begin{aligned}
E(Y) = exp(a0 + a1*P + a2*T)
\end{aligned}$$

$$\begin{aligned}
E(T) = exp(b0 + b1*P)
\end{aligned}$$

To implement the simulation study more easily, we will fix a0, a1, b0 and b1 and vary a2. $a0 = -6; a1 = 1; b0 = -5; b1 = 1$. There will be three treatments available. Therefore, there will be three scenarios: $a2 = -0.5/-1/-1.5$. Also assume that P follows a uniform distribution: $P \sim Unif(0,1)$.

## 2.2 Set seeds
Here, we will generate seeds for 1000 replicates, which will later on be recorded in the output 
```{r}
set.seed(3242)
```

## 2.3 Library
```{r}
library(rms)
library(pec)
library(survival)
library(tidyverse)
library(kableExtra)
```


## 2.4 Function of generating the data
Here is the function of generating the data
```{r}
make.NA <- function(x) {	 
	x[cumsum(cumsum(x))>1] <- NA
	return(x)}


generate <- function(n, time, a0, a1, a2, b0, b1, seeds) {
  # set seeds
  set.seed(seeds)
  # make a matrix
  Y <- matrix(nrow = n, ncol = time)
  treat <- matrix(nrow = n, ncol = time)
  
  # generate the predictors
  P <- runif(n)
  #lambda0 <- exp(a0 + a1*P)/(1+exp(a0 + a1*P)) # hazard without treatment
  #lambda1 <- exp(a0 + a1*P+ a2)/(1+exp(a0 + a1*P + a2)) # hazard with treatment
  lambda0 <- exp(a0 + a1*P) # hazard without treatment
  lambda1 <- exp(a0 + a1*P+ a2) # hazard with treatment
  
  # first time point
  #pT <- exp(b0 + b1*P)/(1+exp(b0 + b1*P))
  pT <- exp(b0 + b1*P)
  treat[,1]		<- rbinom(n,1,pT)
  pY		<- lambda1*treat[,1] + lambda0*(1-treat[,1])
  Y[,1] 	<- rbinom(n,1,pY)
  
  # for time intervals 2 through T:
  for (i in 2:time){
	  treat.t		<- rbinom(n,1,pT)
	  treat[,i]		<- apply(cbind(treat.t, treat[,(i-1)]),1,max)
	  pY		<- lambda1*treat[,i] + lambda0*(1-treat[,i])
	  Y[,i] 	<- rbinom(n,1,pY)
  }
  
  ID 	<- rep(1:n,each=time)
  pred 	<- rep(P,each=time)
  time_point 	<- rep(1:time,n)
  Y.vect 	<- as.vector(apply(Y,1,make.NA))
  treat.vect 	<- as.vector(t(treat))
  start 	<- time_point-1
  stop 		<- time_point
  
  data <- data.frame(ID=ID, 
			  pred=pred,
			  treatment=treat.vect,
			  status=Y.vect,
			  start=start,
			  stop=stop)

# and exclude information after first event:
  dat <- subset(data,subset=is.na(status)==0)

# return a list
  return (dat)
}
```

# 3 Simulation

```{r}
n.sim <- 1000 
n.sce <- 3

n 	<- 100	# no. subjects
time 	<- 50	# no. time intervals
t.pred <- 25	# time point for prediction 


b0 <- rep(-5, 3)
b1 	<- rep(1, 3)
a0 	<- rep(-4.8, 3)
a1 	<- rep(1, 3)
a2 	<- seq(-10,-0.5,length.out = 3)

seeds <- matrix(sample(1:10^8, n.sim*n.sce, replace = F), ncol = n.sce, nrow=n.sim)
output <- matrix(ncol=n.sce, nrow=n.sim)
colnames(output) <- c("treatment1: a2=-1.5", "treatment2: a2=-1", "treatment3: a2=-0.5")

for (i in 1:n.sce) {
  for (j in 1:n.sim) {
    data <- generate(n, time, a0[i], a1[i], a2[i], b0[i], b1[i], seeds[j,i])
    survival <- cph(Surv(start, stop, status) ~ pred, data=data, x = TRUE, y = TRUE, surv = TRUE)
    output[j,i] <- predictSurvProb(survival, newdata = data, times = t.pred)[1]
  }
}

```

# 4 Results

Then summary of the output includes the estimates (mean of the replicates), Monte Carlo standard error (standard deviation of replicates), 2.5% of the distribution of the estimates and 97.5% of the distribution.
```{r}
output.sum <- data.frame(situation = colnames(output)) %>% 
  mutate(estimate = colMeans(output)) %>% 
  mutate(Monte_carlo_se = apply(output, 2, sd)) %>% 
  mutate(lower_bound_se = apply(output, 2, function(x) {quantile(x,probs = 0.025)})) %>% 
  mutate(upper_bound_se = apply(output, 2, function(x) {quantile(x,probs = 0.975)}))

kable(output.sum, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","bordered"),
                full_width = F, row_label_position = "c")
```

Apprently, with a more effective treatment (i.e., lower a2), the survival probability increases but the magnitude of the increase is not much (i.e., less than 1%). Therefore, the treatment usage will not influence the result in prediciton models dramatically.

