---
title: "Assignment 1"
author: "Zhenwei Yang"
date: "2020/9/16"
output: html_document
---


# Perform a small simulation that does the following:

a. Sample 100 samples from a standard normal distribution.

b. For each of these samples, calculate the following statistics for the mean:

*absolute bias

*standard error

*lower bound of the 95% confidence interval

*upper bound of the 95% confidence interval

c. Create a plot that demonstrates the following:

` “A replication of the procedure that generates a 95% confidence interval that is centered around the `
`sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)`

Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.

1. load library 
```{r}
library(tidyverse)
library(ggplot2)
```

2. create seeds and samples
```{r}
set.seed(524)
sim_res <- replicate(100, sample(rnorm(100),100),simplify = F)
```

3. summarize in the table
```{r}
bias <- sim_res %>% 
  lapply(mean) %>% unlist

abs_bias <- abs(bias)
  
se <- sim_res %>% 
  lapply(sd) %>% unlist / sqrt(100)

lower_bound <- bias - 1.96 * se

upper_bound <- bias + 1.96 * se

tbl <- data.frame(sample = 1:100,
                  bias,
                  abs_bias,
                  se,
                  lower_bound,
                  upper_bound,
                  cover = as.factor(lower_bound < 0 & upper_bound > 0))
colnames(tbl) <- c("sample ID", "Mean", "absolute bias", "standard error", "95%CI lower bound", "95%CI upper bound", "Cover")
```

4. plot
```{r}
ggplot(tbl,aes(y = Mean, x = `sample ID`, color = Cover)) + 
  geom_point(aes(color = tbl$Cover)) +
  geom_hline(aes(yintercept = 0),color = "gray", size =2) +
  geom_pointrange( aes(ymin = tbl$`95%CI lower bound`, ymax = tbl$`95%CI upper bound`, color=Cover))
```

5. yield a table
```{r}
tbl %>%
  filter(Cover == F)
```

