---
title: "Synthetic data"
author: "Thom Volker"
date: "12/3/2020"
output: 
  ioslides_presentation:
    logo: MICE_sticker_SMALL.png
bibliography: ref.bib
csl: "apa-6th-edition.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
library(mice)
library(tidyverse)
```

## Real data table

```{r}
dat <- boys %>% mice(m = 1, print = F) %>% complete()

dat %>% DT::datatable(options = list(pageLength = 7))
```

## Real data plot

```{r, fig.align = 'center'}

(real_plot <- 
  dat %>%
    select_if(is.numeric) %>%
    pivot_longer(cols = everything(), names_to = c("Variable")) %>%
    ggplot(., mapping = aes(x = value)) +
      geom_histogram(fill = "#E41A1C", bins = 30) +
      facet_wrap(~ Variable, nrow = 3, ncol = 2, scales = "free") +
      theme_classic() +
      scale_fill_brewer(palette = "Set1") +
      theme(legend.position = "bottom"))
```

## Synthetic data

Synthetic data can be generated easily, like this:

```{r, cache = TRUE, echo = TRUE}
# specify cart and alter predictor matrix
cart <- rep("cart", ncol(dat))
names(cart) <- colnames(dat)
cart['bmi'] <- "~I(wgt / (hgt/100)^2)"

pred <- make.predictorMatrix(dat)
pred[c("wgt", "hgt"), "bmi"] <- 0

syns <- dat %>% mice(m = 5, 
                     method = cart,
                     predictorMatrix = pred,
                     where = matrix(TRUE, nrow(dat), ncol(dat)),
                     print = F,
                     seed = 123)
```

## Real and synthetic data

```{r}
completed <- complete(syns, action = "all")

(comparison_plot <- 
    completed %>%
    map_dfr(function(x) x) %>%
    bind_rows(., dat, .id = "Imputed") %>%
    mutate(Imputed = as.numeric(Imputed)) %>%
    select_if(is.numeric) %>%
    pivot_longer(cols = -"Imputed",
                 names_to = c("Variable")) %>%
    mutate(Imputed = recode(Imputed, `1` = "Synthetic data", `2` = "Observed data")) %>%
    ggplot(., mapping = aes(x = value, fill = Imputed)) +
      geom_histogram(aes(y = ..density..), position = "dodge2", bins = 30) +
      facet_wrap(~ Variable, nrow = 3, ncol = 2, scales = "free") +
      theme_classic() +
      scale_fill_brewer(palette = "Set1") +
      theme(legend.position = "bottom"))
```

## Something interactive

<div style="float: left; width: 30%;">

On the right, there is a plot of the distribution of age, with the actually observed data in red, and the synthetic data averaged over the five imputation rounds in blue.

</div>

<div style="float: right; width: 70%;">


```{r, out.width="100%"}
age_plot <- 
  completed %>%
    map_dfr(function(x) x) %>%
    bind_rows(., dat, .id = "Imputed") %>%
    mutate(Imputed = as.numeric(Imputed)) %>%
    select(age, Imputed) %>%
    mutate(Imputed = recode(Imputed, `1` = "Synthetic data", `2` = "Observed data")) %>%
    ggplot(., mapping = aes(x = age, fill = Imputed)) +
      geom_histogram(aes(y = ..density..), position = "dodge2", bins = 20) +
      theme_classic() +
      scale_fill_brewer(palette = "Set1") +
      theme(legend.position = "bottom")

plotly::ggplotly(age_plot)
```
</div>

## Inferences from synthetic data

To make correct inferences from the synthetic data, we need to use the correct estimators. For instance, we could have

<div style="float: left; width: 50%;">

$$
\begin{align}
\bar{q}_m &= \frac{1}{m}\sum^m_{i=1} q^{(i)}, \\
b_m &= \sum^m_{i = 1} \frac{(q^{(i)} - \bar{q}_m)^2}{m-1},
\end{align}
$$

</div>

<div style="float: right; width: 50%;">

$$
\begin{align}
\bar{u}_m &= \frac{1}{m} \sum^m_{i = 1} u^{(i)}, \\
T_f &= (1 + \frac{1}{m})b_m - \bar{u}_m,
\end{align}
$$
</div>

\ 
\

with $\bar{q}_m$ the mean of the estimates, $b_m$ the between-variability, $\bar{u}_m$ the within data variability, $u^{(i)}$ the within-variance and $T_f$ the total variance of the estimate, as proposed by @raghu.

## References


