---
title: "Markup: Exercise 2"
author: Zo&euml; Dunias
date: "26-10-2020"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
    number_sections: true
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 20px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 16px;
}
h3 { /* Header 3 */
  font-size: 14px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>

---

# Aim
The aim is to verify that for the Poisson distribution $\mu = \lambda$, and to show the relation between skewness and kurtosis and the parameter $\lambda$ using simulation.    

---

# Set up
In this section, the seed is set and the required packages are loaded. 

---

## Set random seed
For reproducibility, fix the seed. 

```{r}
set.seed(100)
```

---

## Load packages

```{r}
library(e1071)
```

---

# Poisson distribution and its mean
In order to verify that for the Poisson distribution $\mu = \lambda$, 1000 samples *(n = 100)* are drawn from a Poisson distribution with $\lambda = 0.3$.
The mean of these samples are calculated and plotted.


```{r}
lamb = 0.3

p <- replicate(10000, rpois(n = 100, lambda = lamb))

M <- colMeans(p)

{hist(M, 
    main = expression(paste('Distribution of sample means (Poisson ', lambda, ' of 0.3)')),   
    xlab = 'sample mean',  
    nclass=20) 
abline(v = lamb, lty = 1, col = "red") # Add the lambda value to the plot as a red line.
}
```

As expected, the distribution of sample means is centered around the value of $\lambda$.

---

# Simulate Poisson distributions with varying $\lambda$ values  
The (a)symmetry of the Poisson distribution depends on the parameter $\lambda$. As $\lambda$ increases, the Poisson distribution gets more symmetrical.
Samples are drawn from Poisson distributions with varying $\lambda$ values.


```{r}
# The logarithmic distribution in the range of lambda's from 0.1 to 30 
lamb1 <- seq(log10(0.1), log10(30), 0.1)
lamb2 <- 10^lamb1

# Allocate memory
s <- matrix(0, nrow = 1000, ncol = length(lamb2)) 

for (i in 1:length(lamb2)) { 
  s[,i] <- rpois(n = 1000, lambda = lamb2[i])
}

{par(mfrow=c(1,2))
hist(s[,16],
    main = expression(paste('Poisson Distribution with ', lambda, ' of 3')),   
    xlab = 'x',  
    nclass=20)
hist(s[,length(lamb2)],
    main = expression(paste('Poisson Distribution with ', lambda, ' of 25')),   
    xlab = 'x',  
    nclass=20)
}
```

As expected, the distribution of a sample corresponding to $\lambda \approx 3$ is more skewed than the distribution of a sample corresponding to $\lambda \approx 25$. 

---

## Skewness and kurtosis
The shape of a distribution can be described with the skewness and kurtosis.
For the Poisson distribution, skewness is related to $\lambda$ according to the following: $$skewness = \frac{1}{\sqrt(\lambda)}$$

For the Poisson distribution, kurtosis is related to $\lambda$ according to the following: $$kurtosis = \frac{1}{\lambda}$$

The skewness and kurtosis are calculated for the samples with varying $\lambda$ values previously drawn.

```{r}
# Allocate memory
skew <- rep(0, length(lamb2)) 
kurt <- rep(0, length(lamb2)) 

for (i in 1:length(lamb2)) { 
  kurt[i] <- kurtosis(s[,i]) # Calculate kurtosis                
  skew[i] <- skewness(s[,i]) # Calculate skewness
}

{par(mfrow=c(1,2))
plot(1/sqrt(lamb2), skew,
    main = 'Skewness Poisson',   
    xlab = expression(paste(1/sqrt(lambda))), 
    ylab = 'skewness'
    )
abline(0, 1, col = "red")
grid()

plot(1/lamb2, kurt,
    main = 'Kurtosis Poisson',   
    xlab = expression(paste(1/lambda)), 
    ylab = 'Kurtosis')
abline(0, 1, col = "red")
grid()
} 
```

The linear relationships shown, confirm the aforementioned equations.  

---

# Replication
In this section, the results of the study are replicated.

---

## Use a different seed

```{r}
set.seed(101)
```

---

## Poisson distribution and its mean
Again, 1000 samples are drawn from a Poisson distribution with $\lambda = 0.3$.
The mean of these samples are calculated and plotted.


```{r}
p_rep <- replicate(10000, rpois(n = 100, lambda = lamb))

M_rep <- colMeans(p_rep)

{hist(M_rep, 
    main = expression(paste('Distribution of sample means (Poisson ', lambda, ' of 0.3)')),   
    xlab = 'sample mean',  
    nclass=20) 

abline(v = lamb, lty = 1, col = "red") # Add the lambda value to the plot as a red line.
}
```

This results in a slightly different distribution, but the distribution of sample means is again centered around the value of $\lambda$.

---

## Simulate Poisson distributions with varying $\lambda$ values  
Again, samples are drawn from Poisson distributions with varying $\lambda$ values.


```{r}
for (i in 1:length(lamb2)) { 
  s[,i] <- rpois(n = 1000, lambda = lamb2[i])
}

{par(mfrow=c(1,2))
hist(s[,16],
    main = expression(paste('Poisson Distribution with ', lambda, ' of 3')),   
    xlab = 'x',  
    nclass=20)
hist(s[,length(lamb2)],
    main = expression(paste('Poisson Distribution with ', lambda, ' of 25')),   
    xlab = 'x',  
    nclass=20)
}
```

This results in slightly different distributions, but the distribution of a sample corresponding to $\lambda \approx 3$ is again more skewed than the distribution of a sample corresponding to $\lambda \approx 25$.

---

## Skewness and kurtosis
The skewness and kurtosis are calculated for the samples with varying $\lambda$ values previously drawn.

```{r}
for (i in 1:length(lamb2)) { 
  
  s[,i] <- rpois(n = 1000, lambda = lamb2[i])
  
  kurt[i] <- kurtosis(s[,i]) # Calculate kurtosis                
  skew[i] <- skewness(s[,i]) # Calculate skewness
}

{par(mfrow=c(1,2))
plot(1/sqrt(lamb2), skew,
    main = 'Skewness Poisson',   
    xlab = expression(paste(1/sqrt(lambda))), 
    ylab = 'skewness'
    )
abline(0, 1, col = "red")
grid()

plot(1/lamb2, kurt,
    main = 'Kurtosis Poisson',   
    xlab = expression(paste(1/lambda)), 
    ylab = 'Kurtosis')
abline(0, 1, col = "red")
grid()
} 
```

This results in slightly different plots (the skewness and kurtosis values are slightly different for the new samples), but again the linear relationships are shown.  


