---
title: "Exercise 2"
author: "Dandan Tang"
date: "2020/11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1 Aim
In this document we explore the confidence validity of bootstrap estimates. The goal of this investigation is to study the possibility of using the bootstrap to simulate a monte carlo experiment on a single (small) data set.

2 Set up
This section covers the global parameters and the necessary packages for executing the simulation study.

2.1 Fixing the random seed
```{r}
set.seed(202)
```
2.2 Loading the required packages
```{r}
library(bain) # population data from bain 
library(lavaan)  # SEM
library(purrr)    # Functional programming
library(psych)    # Descriptives
library(knitr)    # 
library(kableExtra)# output tables
```
2.3 Number of simulations
```{r}
nsim <-  100
```
3 the population data from bain 
```{r}
sesamesim 
m <- length(sesamesim[,1]) # the sample size 
```
3.1 Confirmatory factor analysis(CFA) Model of interest, this model has one factor with six indicators. 
```{r}
model3 <- 'A  =~ Ab + Al + Af + An + Ar + Ac '
truefit <- sem(model3, data = sesamesim, std.lv = TRUE)
```
4 Draw the bootstrap sets
We use the replicate() function to draw nsim =100 bootstrap sets from the m rows in the sesamesim.
```{r}
simdata <- replicate(nsim, 
                     sesamesim[sample(1:m, m, replace = TRUE), ], 
                     simplify = FALSE)
```
5 Bias
First we evaluate the model of interest on each of the drawn bootstrap samples. We do this with the purrr::map() function, which maps the evaluation on each of the listed elements in the simdata object.
```{r}
fit <- simdata %>% 
  map(~ sem(model3, data = .x, std.lv = TRUE))
```
Then, we extract the estimates from the fit.
```{r}
estimates <- fit %>% 
  map(coef) %>% 
  do.call(rbind, args = .) # bind rows into matrix
```
The obtained average estimates are
```{r}
estimates %>% 
  describe(quant = c(.025, .975)) %>%
  .[, c(2:4, 8:9, 11:12, 14:15)] %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = F)
```
average estimates
```{r}
estimates %>% 
  colMeans
```
6 Confidence Validity
 manually calculated CI’s
For unknown σ2β, the 1−α confidence interval for β is defined as
μ=β^±tn−1,a/2SN−−√.
. The following code extracts the estimates and the standard error
```{r}
manual <- fit %>% 
  map(function(x) cbind(vcov(x) %>% diag %>% sqrt(.), coef(x))) %>% # est and resp. vars
  map(function(x) cbind(x[, 2] - qt(.975, 747) * x[, 1],  x[, 2] + qt(.975, 747) * x[, 1])) %>%
  map(function(x) x[, 1] <= coef(truefit) & coef(truefit) <= x[, 2]) %>%
  do.call(rbind, args = .) # bind rows into matrix
```
7 Results
Below are the coverages for each of the investigated scenarios. Results are depicted for the coverage of the confidence intervals over the CFA estimates for 100 simulations.
```{r}

manual%>% 
  colMeans

```
