---
title: "Random presentation"
author: "Rianne Kraakman"
date: "19-1-2021"
output: ioslides_presentation
logo: uulogo.png
bibliography: references.bib
---

```{r setup, include=FALSE}
library(DT)
library(plotly)
```

## A centered still figure

<center>
  <img src = "https://ca-times.brightspotcdn.com/dims4/default/0716824/2147483647/strip/true/crop/466x600+0+0/resize/840x1082!/quality/90/?url=https%3A%2F%2Fcalifornia-times-brightspot.s3.amazonaws.com%2F04%2F43%2F6a2299ab7b1ef9983bd2801fed13%2Fla-et-st-elmo-and-cookie-monster-to-star-in-fu-001" width = 40%>
  </center>

## An interactive table

```{r}
datatable(iris, options = list(pageLength = 6))
```

## An interactive figure

```{r cache = TRUE, echo = FALSE}
plot <- ggplot(data=iris,aes(x=Sepal.Width, y=Sepal.Length,color=Species)) + geom_point() +geom_smooth() + theme_minimal()

ggplotly(plot)
```

## A 2-column slide

<div class="columns-2">
  Here on the left side I can write some text, but on the other side I have some bullet points.
  The references for the information given can be found on the last slide.

  - The earth is round (or not) [@cohen1994earth]
  - Climate change affects ecology [@thuiller2007climate]
  - Sedentary behavior is a problem [@owen2010too]
</div>

## An aligned multi-row equation

$$
\begin{align}
a &= b + c^2, \\
a - c^2 &= b, \\
\text{left side} &= \text{right side}, \\
\text{left side} + \text{something} &\geq \text{right side}, 
\end{align}
$$

or we could do it like this:
$$
\begin{align}
a = b + c^2, \\
a - c^2 = b, \\
\text{left side} = \text{right side}, \\
\text{left side} + \text{something} \geq \text{right side}, 
\end{align}
$$

## R-code displayed but not executed
```{r eval = FALSE}
summary(iris)
```

## References

