---
title: "Monte Carlo simulation exercise"
author: "Piet Jonker"
date: "9/29/2020"
output: html_document
---

Perform a small simulation that does the following:

a. Sample 100 samples from a standard normal distribution.

```{r simulation function, results=FALSE}

# set seed for reproducable results
set.seed(28)

# simulation function
magicalSampleSimulator <- function(sample_size, simulations_amount){
  
  # create empty vector 
  simulationData <- matrix(data = NA, nrow = sample_size, ncol = simulations_amount)
  
  # perform simulation 
  for (i in 1:simulations_amount){
    
    # sample sample_size times from standard normal distribution and save in a
    a <- rnorm(sample_size)
    
    # put a in ith column of simulationData
    simulationData[,i] <- a
  }
 
  return(simulationData) 
  
}

sample_size        <- 100
simulations_amount <- 500
simulation         <- magicalSampleSimulator(
  sample_size = 100, 
  simulations_amount = 500
  )
```


b. For each of these samples, calculate the following statistics for the mean: 
* absolute bias
* standard error
* lower bound of the 95% confidence interval
* upper bound of the 95% confidence interval

```{r}

means    <- colMeans(simulation)
bias     <- abs(0 - means)
st_error <- sqrt(diag(var(simulation)))/sample_size
ci_upper <- means + 1.96 * st_error
ci_lower <- means - 1.96 * st_error
```


c. Create a plot that demonstrates the following:
“A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)

d. Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.