---
html_document: default
pdf_document: default
title: "Some curves"
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 20px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 16px;
}
h3 { /* Header 3 */
  font-size: 14px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>

---

# Aim

I will fit some curves to simulated data. Two different seeds are used.


---

---

## Model

The outcome y is defined as follows:

$$y = x + 8 *  sin(\frac{x}{2}) + \epsilon$$


---

## Loading ggplot

```{r loading packages}

library(ggplot2)

```


## Simualting the data and fitting some curves

```{r Simulating data and fitting curves}


set.seed(123) # set the seed
n <- 100 # sample size
error <- rnorm(n)
polynomial_orders <- c(1,2,5,7,9,15)

## now the data is simulated.

#data

x_dat <-runif(n, min = -6, max = 6)
y_dat <- x_dat+8*sin(x_dat/2)+ error
Dat <-data.frame(x_dat, y_dat)

# storage matrices

fitted_dat <- matrix(0,nrow = n, ncol = length(polynomial_orders))

### now I will fit some curves


for(i in 1:length(polynomial_orders)) {
  model <-lm(y_dat ~ poly(x_dat, degree = polynomial_orders[i]), data = Dat)
  fitted_dat[,i] <-model[["fitted.values"]]

}

ggplot(Dat, aes(x=x_dat, y=y_dat)) + geom_point()+
  geom_line(aes(y = fitted_dat[,1]), colour = "red")+
  geom_line(aes(y = fitted_dat[,2]), colour = "blue")+
  geom_line(aes(y = fitted_dat[,6]), colour = "green") + 
    geom_line(aes(y = fitted_dat[,4]), colour = "black")+
  xlab("x") + ylab("y")
```

## Let's replicate with a different seed

```{r Replicate with a different seed}

set.seed(999) # set the seed
n <- 100 # sample size
error <- rnorm(n)
polynomial_orders <- c(1,2,5,7,9,15)

## now the data is simulated.

#data

x_dat <-runif(n, min = -5, max = 5)
y_dat <- x_dat+8*sin(x_dat/2)+ error
Dat <-data.frame(x_dat, y_dat)

# storage matrices

fitted_dat <- matrix(0,nrow = n, ncol = length(polynomial_orders))

### now I will fit some curves


for(i in 1:length(polynomial_orders)) {
  model <-lm(y_dat ~ poly(x_dat, degree = polynomial_orders[i]), data = Dat)
  fitted_dat[,i] <-model[["fitted.values"]]

}

ggplot(Dat, aes(x=x_dat, y=y_dat)) + geom_point()+
  geom_line(aes(y = fitted_dat[,1]), colour = "red")+
  geom_line(aes(y = fitted_dat[,2]), colour = "blue")+
  geom_line(aes(y = fitted_dat[,6]), colour = "green") + 
    geom_line(aes(y = fitted_dat[,4]), colour = "black")+
  xlab("x") + ylab("y")
```


## Finally, the session info

```{r}
sessionInfo()
```

