---
title: "Markup languages - Week 2"
author: "Bart-Jan Boverhof"
date: "10/12/2020"
output: html_document
---


The subsequent code helps visualizing bias due to RNG. The function randomly generates a string of *n* values accoring to given probability of *prob*. This is replicated *replications* amount of times. The function outputs the error due to sampling as compared with the true probability *prob*. 

```{r}
binomial.prob <- function(prob, n = 10000, replications = 10000)
{
  y <- matrix(nrow = replications, ncol = n)
  z <-  matrix(nrow = replications, ncol =1)
  
  prob.inv <- (1/prob)
  
  for(i in 1:replications) {
      y[i,] <- round(runif(n,min=1,max=prob.inv))
      
      test.number <- round(runif(1,max= prob.inv))
      
      z[i,] <- sum(y[i,] == test.number) / n
      
    }
  return(abs (mean(z) - prob))
}

set.seed(3791)
(bias <- binomial.prob(prob = (1/10), n =1000, replications = 1000))
```
as can be discerned by running the function, the resulting average bias due to sampling equals `r bias` in seed `r 3791`. 


Replication: 
```{r}
set.seed(3791)
bias2 <- binomial.prob(prob = (1/10), n =1000, replications = 1000)
bias == bias2
```
As can be discerned, replication yields exactly the same value. 


