---
title: " Interesting Facts about Tennis  "
author: "Tabea Röber"
date: "11/26/2020"
output: 
  ioslides_presentation:
    css: styles.css
    logo: me.png
bibliography: references.bib
---

## Content of this presentation {.smaller}

As you probably know, tennis is a racket sport that is widely popular. In this short presentation I'd like to tell you some basic and hopefully also interesting (fun) facts about tennis. 


<div class="columns-2">

  **I will focus on these points: **

  - What is tennis, how do you play it? 

  - Some data

  - Fun Facts

  - Last but not least
  
  
  
  ![](https://images.unsplash.com/photo-1542144582-1ba00456b5e3?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1013&q=80){width=280}


</div>





## What is tennis, and how do you play it? {.smaller}

<center>

Tennis is a sport that can be played individually against a single oppponent
(called *single*), or with teams of two players each (called *double*).


![](https://images.unsplash.com/flagged/photo-1576972405668-2d020a01cbfa?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1653&q=80){width=500}

</center>



## Some data {.smaller}

The table below displays data on the Finals of competitions worldwide, from 1991-2016. [@datahub_2018]


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
library(stringr)
library(plotly)
```

```{r echo=FALSE, cache=TRUE}
dat <- read.csv("match_scores_1991-2016_unindexed_csv.csv")

dat1 <- dat %>% 
  filter(tourney_round_name == "Finals") %>%
  mutate(year = as.numeric(str_split_fixed(tourney_year_id, "-", 2)[,1])) %>%
  select(year, tourney_slug, winner_name, 
         loser_name, match_score_tiebreaks) %>%
  rename(location = tourney_slug, winner = winner_name, loser = loser_name, 
         match_score = match_score_tiebreaks)
```


```{r echo=FALSE }
datatable(dat1, options = list(pageLength = 8), rownames = NULL)
```

## Some data {.smaller}

If you wanted to look at Semi-Finals, you could use the following code: 

```{r eval=FALSE}
dat1 <- dat %>% 
  filter(tourney_round_name == "Semi-Finals") %>%
  mutate(year = as.numeric(str_split_fixed(tourney_year_id, "-", 2)[,1])) %>%
  select(year, tourney_slug, winner_name, 
         loser_name, match_score_tiebreaks) %>%
  rename(location = tourney_slug, winner = winner_name, loser = loser_name, 
         match_score = match_score_tiebreaks)

datatable(dat1, options = list(pageLength = 8), rownames = NULL)
```

This code would produce the exact same table as on the slide before, but for Semi-Finals instead of Finals. 


## Some data {.smaller}

Below, I have visualised Roger Federer's wins over the years. 

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
dat2 <- dat1 %>% 
  group_by(year) %>%
  summarise(Federer = sum(winner=="Roger Federer"), 
            Nadal = sum(winner=="Rafael Nadal")) %>%
  pivot_longer(., cols = c(Federer, Nadal), names_to ="who", values_to ="wins")

p <- dat2 %>%
  ggplot(aes(x=year, y=wins, fill=who)) + 
  geom_bar(stat="identity", position=position_dodge()) + 
  theme_minimal() + 
  scale_fill_brewer(palette="Blues")

ggplotly(p)
```


## Fun Facts [@factfile_2020] {.smaller}

- Every year, **24 tons of strawberries** are consumed during Wimbledon.

- Tennis balls were originally white. In 1986, yellow balls were first introduced at Wimbledon.

- An estimated 54,250 tennis balls are used during Wimbledon.

- The first set of sisters to ever win Olympic gold medals in tennis were Venus and Serena Williams.

## Last but not least {.smaller}

Since I have not yet included a mathematical formula, I will do this now. 

\[\begin{align}
  a &= b + c^2 \\
  a - c^2 &= b, \\
  \text{left side} &= \text{right side,} \\
  \text{left side} + \text{something} &\geq \text{right side,}
\end{align}\]


**Disclaimer: This formula has nothing to do with tennis.**


## References
