---
title: "Exercise 1"
author: "Dandan Tang"
date: "2020/11/20"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1.Sample 100 samples from a standard normal distribution.
```{r}
set.seed(202)
simdata <- replicate(100,rnorm(5000,0,1))
```
2.absoult bias
```{r}
Mean <- apply(simdata,2,mean)
bias <- abs(Mean - 0)
```
3.standard error
```{r}
sdd <- apply(simdata,2,var)
Std.Err <- sdd/sqrt(5000)

```
4.lower bound of the 95% confidence interval and upper bound of the 95% confidence interval
```{r}
con <- qt(.975,4999)*Std.Err
ci_lower <- Mean - con
ci_upper <- Mean + con
```
5.arrange the results
```{r}
result <- data.frame(Mean,bias,Std.Err,ci_lower,ci_upper,
                     cover = as.factor(ci_lower<0 & ci_upper>0))

```
6.Create a plot that demonstrates the following:“A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)
```{r}
library(ggplot2)
interval <- aes(ymax=ci_upper,ymin=ci_lower)
ggplot(result,aes(y=Mean, x=1:100, colour = cover)) + 
  geom_hline(aes(yintercept = 0), color = "dark grey", size = 2) + 
  geom_pointrange(interval) + 
   xlab("Sample 1-100") +
   ylab("Means and 95% Confidence Intervals")
```
7.Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.
```{r}
library(knitr)
library(kableExtra)
kable(result,format = 'html', caption = 'The results of simulated sample',
        align = 'r',row.names = TRUE, booktabs= TRUE)
```


