---
title: "Exercise 2: Reproducibility"
author: "Sofie van den Brand"
date: "Markup Languages and Reproducible Programming in Statistics"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: true
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

---
# Aim

Generate a random part of an image, to see if you can guess what it is:

# Set-up


Setting the seed, loading the necessary packages and uploading the mystery picture.


## First and foremost 
Fix seed
```{r}
set.seed(678)
```

## Loading the necessary packages
```{r message=FALSE, warning=FALSE}
# If some packages have not been installed, simply uncomment the following line:
# source("installpackages.R")
library(imager)
library(tidyverse)

```

## Loading the mystery picture into the environment
```{r}
im <- load.image("mystery.jpg")
```

# "Random" part of the image
First, random numbers are drawn to indicate which part of the image will be shown:
`xbegin` indicates the first row that will be shown, and `xend` the last. 
Same goes for the `ybegin` and `yend`

```{r}
xbegin <- floor(runif(1, min = 1, max = nrow(im)))
xend <- floor(runif(1, min = xbegin, max = nrow(im)))
ybegin <- floor(runif(1, min = 1, max = ncol(im)))
yend <- floor(runif(1, min = ybegin, max = ncol(im)))

```

Then, only a random part of the image will be presented: 

```{r}
imsub(im, x >= xbegin, x <= xend, y >= ybegin, y <= yend) %>% plot
```


What would this mystery image be? 

# Session info

```{r}
sessionInfo()
```


