---
title: "Presentation"
author: "Ruben van den Goorbergh"
date: "1/22/2021"
output:
   ioslides_presentation:
    logo: logo.png

references:
- id: covid
  title: Effects of COVID on College Students’ Mental Health in the United States
  author:
  - family: Son
    given: C
  - family: Hedge
    given: S
  - family: Smith
    given: A
  - family: Wang
    given: X
  - family: Sasangohar
    given: F

  container-title: Journal of Medical Internet Research
  volume: 22
  URL: 'http://citeproc-js.readthedocs.io'
  DOI: 10.2196/21279
  issue: 9
  publisher: JMIR 
  page: 261-263
  type: article-journal
  issued:
    year: 2020
    month: 9
---
# Corona and my my Social Life: a sad, sad story

## Before Corona..


```{r, echo=FALSE, message=FALSE}
library(tidyverse)
month <- seq.Date(from = as.Date("2019/1/1"), to = as.Date("2019/12/1"), by = "month")
number_social_activity <- c(10,15,16,15,16,21,25,25,20,19,15,20)
dat <- tibble(month, number_social_activity)  
dat %>% 
  ggplot(aes(x = month, y = number_social_activity)) +
  geom_col()


```

## All those fun conversations I had...
```{r, echo=FALSE, message=FALSE}
require(DT)
date <- seq.Date(from = as.Date("2019/1/1"), to = as.Date("2019/12/31"), by = "day")
conversations <- round(rnorm(365, mean = 7, sd = 2))

tibble(date, conversations) %>% 
  datatable(options = list(pageLength = 5))
```

## And then Corona came...
```{r, out.width = "70%", fig.align = 'center', echo=FALSE, message=FALSE, cache=TRUE}
library(gganimate)
month <- seq.Date(from = as.Date("2020/1/1"), to = as.Date("2020/12/1"), by = "month")
number_social_activity <- c(16,17,9,0,3,5,6,6,5,3,2,2)
n_month <- seq(1, 12 , 1)
dat2 <- rbind(tibble(month, number_social_activity), dat)  

plot <- dat2 %>% 
  ggplot(aes(x = month, y = number_social_activity)) +
  geom_line()

plot + transition_reveal(month)
```
```{r, eval = FALSE }
a <- 'I'
b <- "am"
c <- "bored"

# a+b+c = ?
```


<div class="columns-2">
## The difference:'(
```{r, echo = FALSE, fig.width=3, fig.height=5}
dat %>% 
  ggplot(aes(x = month, y = number_social_activity)) +
  geom_col()+
  ggtitle("Before")
```
```{r, echo=FALSE, fig.width=3, fig.height=5}
dat2 %>% 
  ggplot(aes(x = month, y = number_social_activity)) +
  geom_col() +
  ggtitle("After")
```
</div>

## Predicting social activity
To predict the number of social interction in the future, I fitted ~~made up~~
this linear combination of predictors: 
$$
\begin{aligned}
soc\_ac\_p\_month =& 15 -10 * lockdown + 3 * spring +  \\ &7 * summer+ 2* fall + -5 * cerfew + \\ & 2 * fight\_with\_girlfriend
\end{aligned}
$$

I'm glad it I don't suffer from serious mental health issues as other students do
accordingly to @covid

## References
