# Method {#method}

## Deterministic versus Random Number Generator

The Random Number Generator will be used to divide tasks among roommates in a pseudo-random manner. This will then be compared to a deterministic approach so that the two methods can be compared. 

First the seed is set so that our results are reproducible. Also, the tidyverse is loaded since we'll need it later.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
set.seed(123)
```

### Deterministic approach {-}

First a function will be created for the deterministic approach. One pattern will repeat itself for the entire sequence of dates specified. We will do this as following:

```{r deterministic schedule}

createSchedule <- function(startDate = Sys.Date(),         # the current date is used as default input
                           endDate   = Sys.Date() + 305,   # 
                           createCSV = "no",                           
                           roommateNames
                           ) {

  # create a string of dates between startDate and endDate
  dateString <- seq(from = startDate,
                    to   = endDate,
                    by   = "days")

  # turn it into dataframe
  data <- as.data.frame(dateString)

  # add weekdays
  data <- data %>%
    mutate(days = weekdays(dateString))

  # filter out fridays
  data <- data %>%
    filter(days == "vrijdag" | days == "friday")

  # shuffle the orders of the name strings so that they can be put into dataframe
  roommateNames2 <- roommateNames[c(5,6,1,2,3,4)]
  roommateNames3 <- roommateNames[c(3,4,5,6,1,2)]

  # add columns with chores
  data <- data %>%
    mutate(gang     = rep_len(roommateNames,  length.out = nrow(data)),
           keuken   = rep_len(roommateNames2, length.out = nrow(data)),
           badkamer = rep_len(roommateNames3, length.out = nrow(data))
           )

  # format dates
  data <- data %>%
    mutate(datum = format(dateString, "%d-%m-%y"))

  # delete weekday and dateString column
  data <- data %>%
    select(datum, 
           gang,
           keuken, 
           badkamer)
  
  if (createCSV == "yes") {
    write.csv(data, file = "schedule.csv")
  } else {
      return(data)
  }
  
}
```

Now, we test the function and show the first 10 rows of the schedule. To do this in a reproducible manner, we specify:

-  a start date
-  an end date
-  the names of my roommates

This will be used throughout this document. Completely by chance, the start date is my birthday and the end date Sinterklaas evening (which is my grandfathers birthday). 

```{r}
startDate     <- as.Date("1996-02-16")
endDate       <- as.Date("1996-12-15") 
roommateNames <- c("thijs", "jur", "wies", "hidde", "piet", "jette")

kable(createSchedule(startDate     = startDate,
                     endDate       = endDate,
                     createCSV     = "no", 
                     roommateNames = roommateNames)[1:10,])
```



### RNG approach 1 {-}

In this version, we use the sample function to sample one name per chore per week, The sample function uses the [RNG](#rng). This approach might not be the most convenient, since it is possible for one person to have all the chores in a given week.

```{r fully random schedule}

createScheduleFull <- function(startDate = Sys.Date(),
                               endDate   = Sys.Date() + 305,
                               createCSV = "no",                           
                               roommateNames
                           ) {

  # create a string of dates between startDate and endDate
  dateString <- seq(from = startDate,
                    to   = endDate,
                    by   = "days")

  # turn it into dataframe
  data <- as.data.frame(dateString)

  # add weekdays
  data <- data %>%
    mutate(days = weekdays(dateString))

  # filter out fridays
  data <- data %>%
    filter(days == "vrijdag" | days == "friday")
  
  data <- data %>% 
    rowwise() %>% 
    mutate(gang     = sample(roommateNames, 1, replace = TRUE),
           keuken   = sample(roommateNames, 1, replace = TRUE),
           badkamer = sample(roommateNames, 1, replace = TRUE)
           )
  
  # format dates
  data <- data %>%
    mutate(datum = format(dateString, "%d-%m-%y"))

  # delete weekday column and rename
  data <- data %>%
    select(datum, 
           gang,
           keuken, 
           badkamer)
  
  if (createCSV == "yes") {
    write.csv(data, file = "schedule.csv")
  } else {
      return(data)
  }
  
}
```

Again, we test the function and show the first 10 rows

```{r}
kable(createScheduleFull(startDate     = startDate,
                     endDate       = endDate,
                     createCSV     = "no", 
                     roommateNames = roommateNames)[1:10,])
```


### RNG approach 2 {-}

Now, we create the another function which uses the sample function to pick roommates for chores. Every other week three names from the string of six names will be randomly assigned for a task. The next week, the three remaining names from that sample will be assigned. This way, a random sample is taken, but we avoid the problem of someone having multiple chores in one week.


```{r semi random schedule}
createScheduleSemi <- function(startDate = Sys.Date(),
                           endDate   = Sys.Date() + 305,
                           createCSV = "no",                           
                           roommateNames
                           ) {

  # create a string of dates between startDate and endDate
  dateString <- seq(from = startDate,
                    to   = endDate,
                    by   = "days")

  # turn it into dataframe
  data <- as.data.frame(dateString)

  # add weekdays
  data <- data %>%
    mutate(days = weekdays(dateString))

  # filter out fridays
  data <- data %>%
    filter(days == "vrijdag" | days == "friday")

  # draw a random sample, fill the first row with the first three elements of the sampled vector and the row after with the remaining three elements of the sampled vector
 for(i in seq(from = 1, to = nrow(data), by = 2)){
    sample       <- sample(roommateNames, 6)
    data[i,3]    <- sample[1]
    data[i,4]    <- sample[2]
    data[i,5]    <- sample[3]
    data[i+1, 3] <- sample[4]
    data[i+1, 4] <- sample[5]
    data[i+1, 5] <- sample[6]
 }
  
  # rename columns for clarity
 data <- data %>% 
   rename(gang     = V3,
          keuken   = V4,
          badkamer = V5)
 
  # format dates
  data <- data %>%
    mutate(datum = format(dateString, "%d-%m-%y"))

  # delete weekday column and rename
  data <- data %>%
    select(datum, 
           gang,
           keuken, 
           badkamer)
  
  if (createCSV == "yes") {
    write.csv(data, file = "schedule.csv")
  } else {
      return(data)
  }
  
}
```

Again, we test the function and show the first 10 rows

```{r}
kable(createScheduleSemi(startDate     = startDate,
                         endDate       = endDate,
                         createCSV     = "no", 
                         roommateNames = roommateNames)[1:10,])
```

We have now showed three methods to create schedules, where two use a RNG. I would really like to do an analysis go into depth on the statistics of the approaches such as double or triple assignments in a week. Also, I could have done some form of simulation to see if the approaches are fair when done X amount of times.

Unfortunetaly, most of my time went into creating the bookdown/markdown format, so I'll save this part for a later endeavour. Thanks for reading!
