---
title: "Markup Languages and Reproducible Programming in Statistics"
subtitle: Week 2 Exercise
author:
    - "[Ihnwhi Heo](https://ihnwhiheo.github.io/)"
    - Department of Methodology and Statistics, Utrecht University
date: 'Last modified: `r Sys.setlocale("LC_TIME", "English"); format(Sys.time(), "%d
  %B %Y")`'
output:
  html_document:
    keep_md: true
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Does something that requires RNG

Imagine Ihnwhi wants to travel to the United States. Since he is open to all states, he decides to let R give him suggestions concerning where to go. To that end, Ihnwhi lets R to name five states.

```{r}
# Let R name random five states in the USA.
sample(state.name, 5)
```
## 2. Fixes the RNG seed

However, Ihnwhi finds out that R is a bit capricious (i.e., grillig in Nederlands). Therefore, he decides to give R a seed value not to let it be capricious.

If you can guess the meaning of the seed value below, contact [`Ihnwhi`](mailto:ihnwhi.heo@gmail.com)! He will send you a present.` I am quite sure nobody will see this statement, though!

```{r}
# Set a seed value for reproducibility
set.seed(322)

# Let R name random five states in the USA.
sample(state.name, 5)
```

All right, Out of the five suggestions, Ihnwhi determines to travel to California since he loves the beach!

## 3. Replicates the results

Ihnwhi, a curious guy, changes his mind. He decides to make a data frame that contains R's suggestions by iterating 100 times.

```{r}
# Set a seed value for reproducibility
set.seed(322)

# Create a function that lets R suggest five US states
suggest.ihnwhi.usa.states <- function(N) {
  recommendation <- sample(state.name, N)
  return(recommendation)}

# Create an empty matrix to store R's suggestions
recom.matrix <- matrix(NA, nrow=100, ncol=5)

# Iterate R's suggestion algorithm by 100 times
recom.initial <- replicate(100, suggest.ihnwhi.usa.states(5), simplify = FALSE)
for (i in 1:100) {
  recom.matrix[i,] <- recom.initial[[i]]
}

# The first six rows of the matrix of R's recommendations
head(recom.matrix)
```

## 4. Communicates the info of your session

If you want to communicate with me sincerely, do not see the output below but contact [`Ihnwhi`](mailto:ihnwhi.heo@gmail.com)! He is waiting for your suggestions!

```{r}
# Print the session information
sessionInfo()
```

