---
title: "RNG assignment"
author: "Cassandra"
date: "27-10-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment

I will keep this assignment *quite* short. 


The first important step is to set the seed, to make sure to obtain reproducible results.
```{r set seed}
set.seed(111)
```


The RNG will be used in a simple illustration of the effect of sample size in certainty.

```{r random binom}
sample30 <- mean(rnorm(30, 100, 10))
sample300 <- mean(rnorm(300, 100, 10))

together <- cbind(sample30, sample300)

print(together)

```
In itself, this is not very insightful, and any differences could be attributed to randomness. For this reason, we want to replicate the results many times. Thereby obtaining a sampling dsitrbution. 

```{r replicate}

More30 <- replicate(100,mean(rnorm(30, 100, 10)))
More300 <- replicate(100, mean(rnorm(300, 100, 10)))


p30 <- hist(More30, col = rgb(1,0,0,0.5), xlim = c(90, 110), ylim = c(0, 50), breaks = 20, main = 'Comparison histogram', xlab = 'mean')
p300 <- hist(More300, col = rgb(0,0,1,0.5), add = T, breaks = 10)
legend("topright", c("300", "30"), col=c("blue", "red"), lwd=10)


```

A clear difference in certainty depending on sample size can be seen. 



##Repeat everything with a different set

```{r set seed 2}
set.seed(222)
```


The RNG will be used in a simple illustration of the effect of sample size in certainty.

```{r random binom 2}
sample30 <- mean(rnorm(30, 100, 10))
sample300 <- mean(rnorm(300, 100, 10))

together <- cbind(sample30, sample300)

print(together)

```
In itself, this is not very insightful, and any differences could be attributed to randomness. For this reason, we want to replicate the results many times. Thereby obtaining a sampling distribution. 

```{r replicate 2}

More30 <- replicate(100,mean(rnorm(30, 100, 10)))
More300 <- replicate(100, mean(rnorm(300, 100, 10)))


p30 <- hist(More30, col = rgb(1,0,0,0.5), xlim = c(90, 110), ylim = c(0, 50), breaks = 20, main = 'Comparison histogram', xlab = 'mean')
p300 <- hist(More300, col = rgb(0,0,1,0.5), add = T, breaks = 10)
legend("topright", c("300", "30"), col=c("blue", "red"), lwd=10)


```

Conclusion is the same :)

```{r pressure, echo=FALSE}
sessionInfo()
```

