---
title: "Reproducability & Repiclation"
subtitle: "The product of two normal distributions"
author: "Rianne Kraakman"
date: "27-10-2020"
output: html_document
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 20px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 16px;
}
h3 { /* Header 3 */
  font-size: 14px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>

---

# Aim
We do a simulation study using a random number generator and replicate the results with a different seed. The simulation study shows that the product of two normal random variables does not, in general, have a normal distribution.

---

# Set up
We start by setting the random seed and loading the necessary packages. 

---

## Fixing the random seed
Because random numbers will be used, the feed of the random number generator has to be fixed, so the results will be reproducable. 
```{r}
set.seed(11)
```

---

## Loading the required packages
```{r message=FALSE, warning=FALSE}
library()
```
This simulation study can be done without extra packages!

---

## Simulate samples from a normal distribution
We create 10 samples consisting of 1000 observations from a normal distribution.
```{r}
samples <- replicate(10, rnorm(1000, mean = .5, sd =.5))
```
And we create 10 more samples consisting of 1000 observations from a different normal distribution.
```{r}
samples2 <- replicate(10, rnorm(1000, mean = 1, sd = 1))
```
---

# Show initial distributions
We first need to show that these samples form two normal distributions. 

---

## Distribution sample set 1

```{r}
h <- hist(samples, xlab = "sampled value", main = "Histogram with normal curve")
xfit <- seq(min(samples),max(samples),length=40)
yfit <- dnorm(xfit,mean=mean(samples),sd=sd(samples))
yfit <- yfit*diff(h$mids[1:2])*10000
lines(xfit, yfit, col="blue", lwd=2)
```
This is a normal distribution. 

---

## Distrubution sample set 2

```{r}
h2 <- hist(samples2, xlab = "sampled value", main = "Histogram with normal curve")
xfit2 <- seq(min(samples2),max(samples2),length=40)
yfit2 <- dnorm(xfit2,mean=mean(samples2),sd=sd(samples2))
yfit2 <- yfit2*diff(h2$mids[1:2])*10000
lines(xfit2, yfit2, col="blue", lwd=2)
```
This is a normal distribution

---

# The product of two normal distributions
Now we take the product of the values drawn from the two different normal distributions and show the distribution.

---

## Obtaining the product
Obtain the product of the values drawn from the normal distributions using element by element matrix multiplication. 
```{r}
product_samples <- samples*samples2
```

---

## Show the distribution
Show the distribution of the product of the samples from the two different normal distributions. 
```{r}
h_p <- hist(product_samples, xlab = "sampled value", main = "Histogram with normal curve")
xfit_p <- seq(min(product_samples),max(product_samples),length=40)
yfit_p <- dnorm(xfit_p,mean=mean(product_samples),sd=sd(product_samples))
yfit_p <- yfit_p*diff(h_p$mids[1:2])*10000
lines(xfit_p, yfit_p, col="blue", lwd=2)
```
This is not a (perfectly) normal distrubution.

--- 

# Replicate with only one of the samples
In the section before this, we aggregated all the drawn samples. We can now show that it also holds for only one of the samples.

--- 

## Sample 7
We show this for sample 7 (out of 10)
```{r}
h <- hist(samples[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit <- seq(min(samples[,7]),max(samples[,7]),length=40)
yfit <- dnorm(xfit,mean=mean(samples[,7]),sd=sd(samples[,7]))
yfit <- yfit*diff(h$mids[1:2])*1000
lines(xfit, yfit, col="blue", lwd=2)
```
This is a normal distribution. 

```{r}
h2 <- hist(samples2[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit2 <- seq(min(samples2[,7]),max(samples2[,7]),length=40)
yfit2 <- dnorm(xfit2,mean=mean(samples2[,7]),sd=sd(samples2[,7]))
yfit2 <- yfit2*diff(h2$mids[1:2])*1000
lines(xfit2, yfit2, col="blue", lwd=2)
```
This is a normal distribution

```{r}
h_p <- hist(product_samples[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit_p <- seq(min(product_samples[,7]),max(product_samples[,7]),length=40)
yfit_p <- dnorm(xfit_p,mean=mean(product_samples[,7]),sd=sd(product_samples[,7]))
yfit_p <- yfit_p*diff(h_p$mids[1:2])*1000
lines(xfit_p, yfit_p, col="blue", lwd=2)
```
This is not a (perfectly) normal distrubution.

---

# Replicate with a different seed
We can replicate all of this with a different seed. 

---

## Fixing the random seed

```{r}
set.seed(1155)
```

---

## Simulate samples from a normal distribution
We create 10 samples consisting of 1000 observations from a normal distribution.
```{r}
samples <- replicate(10, rnorm(1000, mean = .5, sd =.5))
```
And we create 10 more samples consisting of 1000 observations from a different normal distribution.
```{r}
samples2 <- replicate(10, rnorm(1000, mean = 1, sd = 1))
```
---

## Show initial distributions
We first need to show that these samples form two normal distributions. 

---

### Distribution sample set 1

```{r}
h <- hist(samples, xlab = "sampled value", main = "Histogram with normal curve")
xfit <- seq(min(samples),max(samples),length=40)
yfit <- dnorm(xfit,mean=mean(samples),sd=sd(samples))
yfit <- yfit*diff(h$mids[1:2])*10000
lines(xfit, yfit, col="blue", lwd=2)
```
This is a normal distribution. 

---

### Distrubution sample set 2

```{r}
h2 <- hist(samples2, xlab = "sampled value", main = "Histogram with normal curve")
xfit2 <- seq(min(samples2),max(samples2),length=40)
yfit2 <- dnorm(xfit2,mean=mean(samples2),sd=sd(samples2))
yfit2 <- yfit2*diff(h2$mids[1:2])*10000
lines(xfit2, yfit2, col="blue", lwd=2)
```
This is a normal distribution.

---

## The product of two normal distributions
Now we take the product of the values drawn from the two different normal distributions and show the distribution.

---

### Obtaining the product
Obtain the product of the values drawn from the normal distributions using element by element matrix multiplication. 
```{r}
product_samples <- samples*samples2
```

---

### Show the distribution
Show the distribution of the product of the samples from the two different normal distributions. 
```{r}
h_p <- hist(product_samples, xlab = "sampled value", main = "Histogram with normal curve")
xfit_p <- seq(min(product_samples),max(product_samples),length=40)
yfit_p <- dnorm(xfit_p,mean=mean(product_samples),sd=sd(product_samples))
yfit_p <- yfit_p*diff(h_p$mids[1:2])*10000
lines(xfit_p, yfit_p, col="blue", lwd=2)
```
This is not a (perfectly) normal distrubution.

--- 

## Replicate with only one of the samples
In the section before this, we aggregated all the drawn samples. We can now show that it also holds for only one of the samples.

--- 

### Sample 7
We show this for sample 7 (out of 10)
```{r}
h <- hist(samples[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit <- seq(min(samples[,7]),max(samples[,7]),length=40)
yfit <- dnorm(xfit,mean=mean(samples[,7]),sd=sd(samples[,7]))
yfit <- yfit*diff(h$mids[1:2])*1000
lines(xfit, yfit, col="blue", lwd=2)
```
This is a normal distribution. 

```{r}
h2 <- hist(samples2[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit2 <- seq(min(samples2[,7]),max(samples2[,7]),length=40)
yfit2 <- dnorm(xfit2,mean=mean(samples2[,7]),sd=sd(samples2[,7]))
yfit2 <- yfit2*diff(h2$mids[1:2])*1000
lines(xfit2, yfit2, col="blue", lwd=2)
```
This is a normal distribution

```{r}
h_p <- hist(product_samples[,7], xlab = "sampled value", main = "Histogram with normal curve")
xfit_p <- seq(min(product_samples[,7]),max(product_samples[,7]),length=40)
yfit_p <- dnorm(xfit_p,mean=mean(product_samples[,7]),sd=sd(product_samples[,7]))
yfit_p <- yfit_p*diff(h_p$mids[1:2])*1000
lines(xfit_p, yfit_p, col="blue", lwd=2)
```
This is not a (perfectly) normal distrubution.

---

END OF DOCUMENT

---
```{r}
sessionInfo()
```
